---
title: "Problem Set 4 Part 2"
format: typst
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This part of the problem set involves the usage of R and the submission of R codes. Please fill in your R codes and answers in the indicated code chunks/areas. When finished, please knit the document to PDF and submit the PDF file to Moodle.

In the case that your document does not knit to PDF, you may try the following:

1.  Clear your local environment
2.  Re-run all your cells in order and pay attention to whether you've defined variables after you called them.
3.  Knit to html and print as PDF.

**You will not receive credit if the markdown fails to compile to PDF. For example:**

- **No PDF file is submitted**

- **PDF of a raw markdown file is submitted**

- **PDF of screenshots is submitted**

**If you rendered with the global** `eval=False` **option, you will receive a 2-point deduction.**

Before we begin, we will need to load the necessary packages:

```{r, message=FALSE}
set.seed(105)
```

\newpage

# Chi-Squared Distribution

In statistics, the Normal distribution is very useful. In fact, other distributions often have a relationship with the Normal distribution. In this case, let's look at the Chi-Squared distribution with 1 degree of freedom, which can be denoted as

$$
X\sim \chi^2_1\,.
$$ We have the relationship

$$
X = Z^2
$$ where $Z\sim N(0,1)$.

## Question 1

Generate a random vector of length 100 from a standard normal distribution and call it `z`. (2pts)

```{r}
## Your code here

```

## Question 2

Now, square `z` to obtain a chi-squared random variable, and call it `x`. Note: you can square something in R by using the `^2` notation. (2pts)

```{r}
## Your code here

```

## Question 3

It turns out that we sample a chi-squared random variable in R directly using `rchisq` function. Use this to sample 100 realizations of $X$ directly and call it `x2` in R. Note that you need to specify `df=1` as an argument in `rchisq`.

```{r}
## Your code here

```

## Question 4

Create a histogram for both `x` and `x2`. You can plot them separately if you would like. Make sure you title your plot so we know which comes from normal and which comes from Chi-Squared. (4pts)

```{r}
## Your code here

```

## Question 5

Are the plots you produced in the previous question similar? (2pts)

**Your Answer**:
