---
title: "Problem Set 3 Part 2: R"
format: typst
---

This part of the problem set involves the usage of R and the submission of R codes. Please fill in your R codes and answers in the indicated code chunks/areas. When finished, please knit the document to PDF and submit the PDF file to Moodle.

In the case that your document does not knit to PDF, you may try the following:

1.  Clear your local environment
2.  Re-run all your cells in order and pay attention to whether you've defined variables after you called them.
3.  Knit to html and print as PDF.

**You will not receive credit if the markdown fails to compile to PDF. For example:**

- **No PDF file is submitted**

- **PDF of a raw markdown file is submitted**

- **PDF of screenshots is submitted**

**If you rendered with the global** `eval=False` **option, you will receive a 2-point deduction.**

Before we begin, we will need to load the necessary packages:

```{r, message=FALSE}
set.seed(105)
```

\newpage

# Binomial Distribution in R

Suppose we are interested in the distribution $X\sim BIN(70, 0.8)$

## Question 1

Find the probability $P(X=65)$. (2pts)

```{r}
## Your code here

```

## Question 2

Find $F_X(30)$. (2pts)

```{r}
## Your code here

```

## Question 3

Estimate the median of $X$ using the CDF (the nearest integer is fine). Don't sample from $X$. (2pts)

```{r}
## Your code here


```

## Question 4

Sample 100 realizations of $x$. Name the output `x1`. (2pts)

```{r}
## Your code here

```

## Question 5

Using the normal approximation technique as shown in Question 6e of Part 1, sample 100 realizations of $X$ from the normal distribution. Note: You need to find the $\mu$ and $\sigma^2$ for this approximation. Name the output vector `x2`. (4pts)

```{r}
## Your code here
rnorm(100, )
```

## Question 6

Create a histogram of `x1` and include a title that includes the distribution where the data are from. (2pts)

```{r}
## Your code here

```

## Question 7

Create a histogram of `x2` and include a title that includes the distribution where the data are from. (2pts)

```{r}
## Your code here

```

## Question 8

What do you observe from **question 6** and **Question 7**? Are the distributions similar? (2pts)

**Your answer**:

# Inverse CDF Sampling

Sometimes, it is nontrivial to sample from a distribution, especially the uncommon ones. One common technique is called the Inverse CDF methods. Suppose

$$
X\sim U(0,1)
$$

and we know the inverse of the CDF of the target distribution $F_Y^{-1}$. Then, we can perform sampling of $Y$ simply by the formula

$$
Y\sim F_Y^{-1}(x)\,.
$$

## Question 9

Sample 1000 realizations of $X$ from $X\sim U(0,1)$, and call it `x` in R. (2pts)

```{r}
## Your code here

```

## Question 10

Use the inverse CDF method to sample 1000 observations from $N(0, 1)$ and call it `y` in R. (2pts)

```{r}
## Your code here

```

## Question 11

Plot a histogram of `y` and give it a title with the method of sampling. (2pts)

```{r}
## Your code here

```

## Question 12

Does the histogram from **Question 11** look normally distributed? (1pts)

**Your answer**:
