---
title: "Problem Set 2: R"
format: typst
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This part of the problem set involves the usage of R and the submission of R codes. Please fill in your R codes and answers in the indicated code chunks/areas. When finished, please knit the document to PDF and submit the PDF file to Moodle.

In the case that your document does not knit to PDF, this is the time to debug and ask me questions before the review. You may try the following:

1.  Clear your local environment
2.  Re-run all your cells in order and pay attention to whether you've defined variables after you called them.

Before we begin, we will need to load the necessary packages:

```{r, message=FALSE}
library(tidyverse)
```

\newpage

# Measure of Dispersion: Mean Absolute Deviation

When we talk about the **dispersion** of a data set we are concerned with how spread out it is. Both standard deviation and IQR are measurements of dispersion. However, they are not the only measurements of dispersion.

One alternative statistic we can use is the **mean absolute deviation**. We can start from

$$\bar X - X_i\,,$$

which is the deviation. Then the absolute deviation is

$$|\bar x - x_i|$$

which takes the absolute value of $\bar x - x_i$. The **mean absolute deviation** is defined by taking the mean, which is

$$\frac{\sum_{i=1}^n |\bar x -x_i|}{n}\,.$$

Suppose we have the following data set:

```{r}
x <- c(2, 0, 6, 28, 19, 6)
```

## Question 1

Write code to find the mean of `x` and save it in a variable called `mean_x`.

```{r}
## Your code goes here

```

## Question 2

Find the vector $\bar x - \boldsymbol x$. (Note $\bar x$ is a scalar number and $\boldsymbol x$ is a vector. In R, you can subtract a scalar from a vector directly. Try `x-1` and see what happens).

```{r}
## Your code goes here

```

## Question 3

Use your codes from the previous question and find the absolute value to obtain $|\bar x - \boldsymbol x|$. In R, you can find the absolute value using the `abs()` function.

```{r}
## Your code goes here

```

## Question 4

Use your codes from the previous question and find

$$\frac{\sum_{i=1}^n |\bar x -x_i|}{n}\,.$$

Note that the form of $\frac{\sum_i y_i}{n}$ is simply the arithematic mean. So, you may consider using the `mean()` function.

```{r}
## Your code goes here


```

# Real Data Application

## Question 5

Load the dataset named `food_access.csv` in the data folder on Jupyter Hub. Make sure you use the link on Moodle so that this notebook and the data are in the correct place. Use the function `read_csv`, not `file.choose()` or `read.csv()`.

Hint: The file path should be `"../data/food_access.csv"`

```{r}
## Your code goes here

```

## Question 6

Filter the data using the `State` variable to include only observations from North Carolina. Name the output dataset `Food` again.

```{r}
## Your code goes here

```

## Question 7

Use your codes from **Question 4** to the find the mean absolute deviation for the `NUMGQTRS` variable. (Note: To extract a column as a vector, use the `df$column` syntax.)

```{r}
## Your code goes here


```

## Question 8

Use the data dictionary for the dataset (named `food_access_dictionary.csv` on Moodle) to determine the meaning of the `NUMGQTRS` variable.

**Your answer**:
