---
title: "Week 1"
output: pdf_document
date: "2026-08-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R

## Hello World

These are the most basic R operations:

```{r}

```

We can also assign values to variables:

```{r}

```

## Collections

What if we have many values that we need to store? We can use `vector`, `list`, or `data.frame`. To start, let's generate a sequence of numbers in a **vector form**:

To generate our own, we can use the following syntax:

```{r}

```

A list can store many vectors or other types of data:

```{r}

```


## Functions and Help

In R, we can call functions to perform tasks (so that we don't have to write everything from scratch)! For example, we can find the sum (i.e. add things together):

```{r}

```

Notice that `c()` is also a function!

To get help, we can use the following syntax:

```{r}

```

## Packages

Let's not reinvent the wheels if we don't have to! That's what packages are for: we can use what others have created. One package we will be using is `tidyverse`, which is very helpful for data analyses.

Let's first install the package:

```{r, eval=FALSE}

```

Then, let's load the package:

```{r}

```


## Reading Datasets

In the class, we will mainly be using CSV (comma-separated values) files, which are widely compatible with R.

Let's read in a dataset we'll be using today:

```{r}

```

## Tidy Data Format

Let's inspect the data again:

```{r}

```

- Each row is an observation (i.e. county)
- Each column is a variable

# Types of Data

## Categorical Data

```{r}

```

Also note *nominal* vs *ordinal* data as introduced in OIS.

## Numerical Data

### Continuous Data

```{r}

```

### Discrete Data

```{r}

```

# Population vs Sample

**Population**:

**Sample**:

**Census**:

## Data Example: US County Data

```{r}
df <- usdata::county
head(df)
```

**Questions**:

- If we want to study the poverty rate across all US counties, what's the population of interest?

- How do we take a sample?

## Detour: Some more R basics

### The Pipe Operator

The pipe operator from Tidyverse takes something and pass it as the first argument to a function!

```{r}

```

This is particularly useful for dataframes.

### The Sample Function in R

We can take a simple random sample:

```{r}

```

## Data Example:

- How do we take a sample?

First, let's do so in Tidyverse:

```{r}

```

```{r}

```

Now, let's do it in base R:

```{r}

```

Each has its advantages and quirks.

# Parameters vs Statistics

**Parameters**:

**Statistics**:

## Data Example

**Question**: We are interested in the county with the highest poverty rate in the US and the corresponding poverty rate. What are our parameters?

```{r}

```

**Question**: What happens when we take a sample? What are our statistics?

```{r}

```

**Question**: Is our statistic close to the parameter?

# Inference, Biases, and Representativeness

We want our statistics to reasonably reflect the parameters of interest for valid **inference**:

Ideally, we want our samples to be representative about the populations of interest.

- **Can we always trust our statistics?**

- **What sources of biases or error do we have?**

# Sampling Methods

## Simple Random Sample (SRS)

**Definition**:

Our previous example on sampling counties is indeed an SRS!

To implement SRS in R, we use the `sample` function:

```{r}

```

This is done without replacement!

## Stratified Sample

**Definition**:

To do it in R:

```{r}

```

Let's verify if it worked:

```{r}

```

## Cluster Sample

**Definition**: We select some of groups from the population.

Example: Let's sample five states from the `county` dataset and examine all counties.

```{r}

```

## Systematic Sample

**Definition**:

This is somewhat complicated to do. We won't do it in R now.

## Convenience Sample

**Definition**:

**Examples**:

- 

- 

# Bonus Contents:

## Loops

Sometimes, we need to perform an operation to every member of a collection.

```{r}

```

Of course, we can do it directly:

```{r}

```

## Indexing Datasets

Sometimes we want to look at some specific rows or columns of a data frame. We can use the `[]` notation to do that. To get the first 10 rows and the first 2 columns:

```{r}

```

We can also index columns by name. Notice that we use vectors to index multiple things:

```{r}

```

We can also specify conditions (filtering):

```{r}

```