---
title: "W2D1"
format: typst
---

```{r setup, include=FALSE, message= F, warning=F}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Mean

Mean is also known as the average.

- **Population mean** is denoted as $\mu$
- **Sample mean** is denoted as $\bar x$

We define the sample mean as the following:

$$\begin{aligned}
\bar x &= \frac{x_1+x_2 + \cdots+x_n}{n} \\
&= \frac{1}{n}\sum_{i=1}^{n} x_i\,. 
\end{aligned}$$

Let's write the definitions writing on the board.

Let's do some examples in R!

```{r}

```

More usefully, we often want to investigate results from a dataset (2017 population from county data):

```{r}

```

Let's try again!

```{r}

```

Let's make it more interesting: Find the average county population by state!!

```{r}

```

# Median

Median can be thought of as the middle number: 50% of data above.

## Two Cases of the Medians

1.  Odd number of observations

We use the middle number.

```{r}

```

2.  Even number of observations

We average the middle two numbers.

```{r}

```

- **Why is median desirable?**

It is unaffected by extreme values. Examples:

```{r}

```

```{r}

```

## County Data Example

```{r}

```

# Histogram

Histogram showcases the shape of one variable.

For example, let's plot a histogram of poverty rates of US counties:

```{r}

```

But this is not as informative as it can be:

```{r}

```

**Question**: Are the observed data symmetric around its median?

```{r}

```

The answer is no!

## Skewness

### Symmetric data

```{r}

```

```{r}

```

### Left Skewed

```{r}

```

```{r}

```

### Right Skewed

```{r}

```

```{r}

```

## Property of a Symmetric Distribution

**Mean and the median are the SAME!!**

```{r}

```

```{r}

```

```{r}

```

**What happens when are not symmetric?**

Let's go back to the county data with an extreme example:

```{r}

```

```{r}

```

```{r}

```

**Why is that?**

## Histogram Bin Widths

### Too Few

```{r}

```

### Too Many

```{r}

```

**Be careful about which you choose! Especially not too few!**

# Scatter Plots

What happens when we want to examine the relationship between 2 variables?

We use **scatter plots**.

Example in R:

```{r}

```

Let's make it better:

```{r}

```

**How do we interpret the relationship?**

# 5 Number Summary in R

**Idea**: Let's look beyond the center of the data!

- How about the minimum and the maximum?
- How about 25% and 75% of the data?

We can easily get more information in R by using the `summary` function on numeric data:

```{r}

```

Let's unpack:

- *Min.*: Minimum, the smallest number
- *1st Qu*: 1st Quartile, 25% of the data
- *Median*: Median, 50% of the data
- *3rd Qu*: 3rd quartile, 75% of the data
- *Max.*: Maximum, the largest
- *NA's*: All missing values

**Caveate**: This doesn't work well with tidyverse!

## Interquartile Range (IQR)

We define IQR as

$$IQR = Q_3-Q_1\,.$$

In R, we can do

```{r}

```

## Problem

Statisticians can't agree on the exact definition of $Q_1$ and $Q_3$.

# Boxplots

Boxplots perfectly encapsulates the five number summary. First, let's create a simple boxplot:

```{r}
x <- data.frame(x = c(1, 2, 2, 4, 7, 10, 15, 17, 20, 21))

```

Let's create the boxplot:

```{r}

```

## Boxplots of Skewed Data

Left Skewed:

```{r}

```

Right skewed:

```{r}

```
