---
title: "W2D2"
format: typst
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
set.seed(105)
```

# Measuing the Spread of Data

We've talked about the mean and median as characteristics of distributions and data we observe. However, the following data have the same mean and median but with different variances:

```{r}
df <- data.frame(x1 = rnorm(1000),
                 x2 = rnorm(1000, sd=2))


```

So, we need to measure how spread out the data are!

# Variance and Standard Deviation

## Variance

Population variance is defined as the following

$$\sigma^2 = \frac{\sum_{i=1}^n (x_i-\mu)^2}{n}$$

However, do we know $\mu$? We typically don't. Instead, we compute the sample variance:

$$s^2 = \frac{\sum_{i=1}^n (x_i-\bar x)^2}{n-1}$$

**Question**: When $n$ is large, does $n-1$ vs $n$ matter?

## Standard Deviation

Standard deviation is simply the square root of the variance formulations:

$$\sigma = \sqrt{\sigma^2} = \sqrt{\frac{\sum_{i=1}^n (x_i-\mu)^2}{n}}$$

and

$$s = \sqrt{s^2} = \sqrt{\frac{\sum_{i=1}^n (x_i-\bar x)^2}{n-1}}$$

## Boxplots of Different Variance

Now, let's plot the boxplots side by side:

```{r}

```

**What do you notice about these plots?**



## Let's do it in R!!

We can compute the sample variance in R using the county data! Specifically, let's look at data from North Carolina, so that we don't need to deal with missing values.

```{r}
df <- usdata::county

```

```{r}

```

And standard deviation is the following:

```{r}

```

```{r}

```

Now, we've verified that they are the same!

**When is variance useful in analyzing real data?**

Suppose we want to find out how spread out populations are across the years.

```{r}

```

### Real Example

Let's create a boxplot of 2007 Population from the county data:

```{r}

```

# Outliers

As observed in the data above, some values can be "relatively extreme", if we compare them to the rest. They are called **outliers**. We need a more rigorous definition. An observation is considered an outlier if it is outside the interval

$$(Q_1-1.5\times IQR,\, Q_3 + 1.5\times IQR)$$

Let's do it in R!

```{r, eval = FALSE}
df %>%
  filter(pop2017 > fivenum(pop2017)[4]+1.5*IQR(pop2017) |
         pop2017 < fivenum(pop2017)[2]-1.5*IQR(pop2017))
```

Similarly, you can use the bracket notations from base R to perform the subset!

\newpage

# Understanding One Categorical Variable

When we have only one categorical variable, we are interested in the counts: namely how many observations are associated with each value. Let's look at `county` dataset with a focus on NC again.

Notice that there is a column called `metro`:

```{r}

```

**Question**: How many counties in the US contain a metropolitan area?

```{r}

```

We can of course do in base R:

```{r}

```

**Question**: What percentage of the counties contains a metro area?

```{r}

```

## More Than 2 Categories

Let's look at the median education for the county!

```{r}

```

```{r}

```

# Two Categorical Variables

For two categorical variables, we look at contingency tables!

## Simple 2 by 2 Case

Using the example from textbook, let's first create a variable that quantifies whether a state

```{r}

```

**Question**: Is there a relationship between population gain and metro area status?

```{r}

```

**Questions**:

1.  How many metro counties gained population?
2.  What percentage of counties that lost population does not contain a metro area?
3.  How many counties lost population?

## Larger Contingency Tables

We can also generalize this to more than 2 by 2:

```{r}

```

The interpretations are the same!

# Visualizing A Categorical Variable

For categorical variables, we usually use bar charts!

## Metro

First, let's plot the bar chart:

```{r}

```

But this plot lacks precise information on counts, unlike 

```{r}

```

## Median Education Level

Now, let's plot the median education, but here is a twist: we need percentages.


```{r}

```
