---
title: "W3D1: R Portion"
format: typst
---

# Loading an External Dataset

It's important for us to load external datasets because not all datasets are built into R.

```{r}

```

Now that we have the dataset loaded, let's do some preprocessing:

```{r}
df <- df %>% 
  mutate(metro = factor(metro, levels = c("no", "yes")),
         median_edu = factor(median_edu, levels = c("below_hs", "hs_diploma", "some_college", "bachelors")),
         smoking_ban = factor(smoking_ban, levels = c("none", "partial", "complete")))
```

In our class, we will be using `csv` files, which are called `comma separate values`. (Note: these files can be opened with MS Excel as well.)

# Two Categorical Variables

For two categorical variables, we look at contingency tables!

## Simple 2 by 2 Case

Using the example from textbook, let's first create a variable that quantifies whether a state

```{r}

```

**Question**: Is there a relationship between population gain and metro area status?

```{r}

```

**Questions**:

1.  How many metro counties gained population?
2.  What percentage of counties that lost population does not contain a metro area?
3.  How many counties lost population?

## Larger Contingency Tables

We can also generalize this to more than 2 by 2:

```{r}

```

The interpretations are the same!

## Visualizing Two Categorical Variables

We can plot a side-by-side bar chart:

```{r}

```

Or a proportion (filled) bar chart:

```{r}

```

Which one you use depends on the message you want to convey.

**Food for thought**: What information does the filled bar chart not convey?

# Visualizing Continuous Variable by Group

It is often the case that we want to investigate a continuous variable by a categorical variable: for example, whether the poverty rate is higher for counties with population gain! To do this, we can visualize the

```{r}

```

**Questions**: 1. What can we conclude from the plot? 2. Are there outliers in this case? 3. Which group has a larger IQR?

### Side by Side Boxplots with More Categories

```{r}

```

# Bad Data Visualization

Looking at the previous plot:

1.  Should we remove the `below_hs` column from the plot?
2.  If we have missing values, should we include them in the plot?

You can try the following code at home:

```{r}

  
```

and see whether we have missing values.

Examples of other bad data visualizations:

