---
title: "W6D1"
format: typst
---

```{r}
set.seed(105)
```

# Uniform Distribution

In R, we have the following functions to sample from a uniform distribution:

- `runif`: Sample random numbers from the specified uniform distribution
- `dunif`: Find the density for at a given $x$
- `punif`: Find the CDF $F(x)$ at a given $x$
- `qunif`: Find $F^{-1}(x)$, where an input of cumulative probability returns $x$

## Sampling from the Uniform Distribution

```{r}

```

Let's check that the data are actually uniform:

```{r}

```

## Find the Density

```{r}

```

## Find the CDF

```{r}

```

## Find the Inverse CDF (Quantile)

```{r}

```

# Plot Densities/PMF in R

## Plot Binomial PMF

Suppose $X\sim BIN(30, 0.2)$. We can first find the PMF in R:

```{r}

```

Then, let's plot the PMF:

```{r}

```

# Normal Distribution

In R, we have the following functions to sample from and find quantities a normal distribution:

- `rnorm`: Sample random numbers from the specified normal distribution
- `dnorm`: Find the density for at a given $x$
- `pnorm`: Find the CDF $F(x)$ at a given $x$
- `qnorm`: Find $F^{-1}(x)$, where an input of cumulative probability returns $x$

# Sampling from the Normal Distribution

Let's sample from the standard normal distribution $N(0, 1)$:

```{r}

```

Let's check that the data are actually normal using an empirical density plot:

```{r}

```

# Find the Density

This is not as useful as finding the probability for a PMF, but we can still do it:

```{r}

```

Let's plot the actual normal density:

```{r}

```

## Find the CDF

For the standard normal distribution, let's verify the empirical rules in the lecture notes.

Recall that we can write

$$
P(a \leq X\leq b) = F(b)-F(a)\,.
$$

For 1 standard deviation:

```{r}

```

For 2 standard deviation:

```{r}

```

For 3 standard deviation:

```{r}

```

## Find the Inverse CDF (Quantile)

Suppose you are interested in finding the $x$ value such that you are in the top 10% of of a standard normal distribution.

```{r}

```

\newpage

# Examples of a Normal Distribution

Let $X$ denote the numbers of grams of hydrocarbons emitted by an automobile per mile. Assume that $X$ is normal with $\mu=1$ grams and $\sigma=0.25$.

## Probability Calculation

What is the probability that a randomly selected car emits between 0.9 and 1.54 grams?

```{r}

```

What is the probability that a given car emits more than 0.7 grams?

```{r}

```

## Z Scores

Suppose a car is measured to emit 1.49 grams. How many standard deviation away is this observation from the mean?

```{r}

```

## Quantile Calculation

A manufacturer wants to design a new car with emission in the 10th percentile. What is the emission requirement needed to achieve that?

```{r}

```

\newpage

# Normal Approximation to the Binomial Distribution

## Aside: For Loop

We use a `for` loop to iterate over a range of values. You can read it "for each `i` in the vector, run the following code."

```{r, eval=FALSE}
for (i in 1:10) {
  print(i)
}
```

## Visualizations of Binomial Distribution

First, let's visualize a series of binomial distributions, each with increasing $n$:

```{r}
par(mfrow = c(2, 3))

for(n in c(5, 10, 20, 30, 50, 100)) {
  plot(x = 0:n, y = dbinom(0:n, n, 0.2),
       type = "h",
       main = paste0("n = ", n),
       ylab = "PMF",
       xlim = c(0, n))
}
```

Now, let's be fancy and add a normal curve on top of each:

```{r}
par(mfrow = c(2, 3))

for(n in c(5, 10, 20, 30, 50, 100)) {
  plot(x = 0:n, y = dbinom(0:n, n, 0.2),
       type = "h", lwd=2,
       main = paste0("n = ", n),
       ylab = "PMF", xlab = "x",
       xlim = c(0, n))
  
  lines(x = seq(0, n, 0.01), y = dnorm(seq(0, n, 0.01), 0.2*n, sqrt(0.2*0.8*n)),
        col = "red", lwd = 2)
}
```

When $n$ is large, we get a very good approximation!

## Example

Suppose you know that the city you live in rains 30% of the days during a given year. Assume that the weather is independent for each day, and randomly decided.

Let $X$ be the number of days that rains, and we have

$$
X\sim BIN(365, 0.3)\,.
$$

### Mean and Variance

Let's find the mean and variance for this case.

```{r}

```

### Probability Calculations

What is the probability that it rains more than 100 days a year?

We want to find

$$
P(X=100)\,.
$$

Using binomial distributions:

```{r}

```

Using Normal approximation:

```{r}

```

**Question**: Are they close? Why are they not exactly the same?
