---
title: "W6D2_R"
format: typst
execute:
    eval: true
---

# Sample Means

```{r}
set.seed(2026)
```

Let's examine sample means closer. We know that suppose

$$
X\sim U(0,1)
$$

the mean is $0.5$. Let's first take a sample and compute its sample mean:

```{r}

```

## Law of Large Numbers

From law of large numbers, we know that sample mean converges to population mean:

$$
\lim_{n\to \infty} \bar x = \mu
$$

Let's use a simple example: we can look at the error when we increase the sample size.

```{r}
abs(mean(runif(2)) - 0.5)
abs(mean(runif(10)) - 0.5)
abs(mean(runif(100)) - 0.5)
abs(mean(runif(1000)) - 0.5)
```

Now, we can be a little fancy. Let's show that it works by repeated sampling and increasing sample size (yes, we are just being fancy and you don't need to reproduce this).

```{r}
# Set up parameters
mu <- 0.5
n <- c(1, 10, 100, 1000, 10000)

# Sample 
error <- list()
counter <- 1
for (i in n) {
  error[[counter]] <- numeric(100)
  for (j in 1:100) {
    error[[counter]][j] <- mean(runif(i)) - mu
  }
  counter <- counter + 1 
}

par(mfrow = c(1, 5))
for (i in 1:5) {
  boxplot(error[[i]], ylim = c(-0.5, 0.5),
          main = paste0("n=", n[i]),
          ylab = "Error")
}
```

We can show that two things happen:

1.  The sample mean is unbiased
2.  As the sample size increases, the variance of the sample mean decreases.

# Visualizing the CLT for Means

**Question**: What would you expect to happen when we repeatedly compute the mean from any distribution?

Let's try it! Again, let's use

$$
X\sim U(0,1)
$$

which is certainly not normally distributed.

For the following code, we need to run it in the console rather than in the note book:

```{r}
# x11()
sample_mean <- numeric(100)

for (i in 1:100) {
  sample_mean[i] <- mean(runif(100))
  # hist(sample_mean[1:i], xlim = c(0, 1),
  #      main = "CLT for Uniform Mean")
  # Sys.sleep(0.1)
}
```

# The Central Limit Theorem

The Central Limit Theorem (CLT) can be stated as

$$
\sqrt{n}\frac{\bar X-\mu}{\sigma} \to N(0,1)
$$

We can alternative express it as (with a bit abuse of notation on the idea of limiting distributions):

$$
\bar X \sim N\left(\mu, \frac{\sigma}{\sqrt{n}} \right)\,.
$$

We'll write down some properties and derivations on paper.

## CLT for Uniform Means

Let's check the mean and variances from the previous example to see whether these are correct. We know that

$$
X \sim U(0, 1)\,.
$$

and we have

$$
\begin{aligned}
\mu &= 0.5\\
\sigma &= \sqrt{\frac{1}{12}}
\end{aligned}
$$

Let's check the actual mean and variance from the sampling distribution

```{r}

```

Let's use R to find the probability $P(0.48\leq \bar X\leq 0.52)$:

```{r}

```

which are very close to what we expect!

## CLT for Skewed Distribution

And in fact, this works for a skewed distribution as well!

```{r}
sample_mean_skewed <- rexp(100*100)
sample_mean_skewed <- matrix(sample_mean_skewed, nrow = 100)
sample_mean_skewed <- colMeans(sample_mean_skewed)
hist(sample_mean_skewed, main = "Sample Means from Skewed Distribution")
```
